---
title: "Network Analysis"
Authours: Lili Benitez, Livia Raulinaitis, Sydney Jansen
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Network Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup}
#library(ADANetworks)
```

## Load libraries, you may need to install the package bipartite
```{r load, message=FALSE, warning=FALSE}
#install.packages("bipartite")
library(tidyverse)
library(bipartite)
hummingbirds<- read.csv("C:/Users/lb37426/Documents/ADA-Collab-Project/EXAMPLE MATRIX_hummingbirdmatrix.csv", header = T) %>%
  column_to_rownames("X")
head(hummingbirds)
```

## Network Visualization: 
Lets look first at what this network looks like: we can do this using a couple of different functions, `plotweb` and `visweb`. The output is a network of lines connecting interacting hummingbird and plant species. 
```{r network visualization}
plotweb(sortweb(hummingbirds),  arrow="both", y.width.low=0.05, labsize=.7, text.rot=90, col.high="blue", 
	col.low="darkgreen")
visweb(as.matrix(hummingbirds), labsize = 1)
```

## Calculating network metrics. 
Now that we have visualized the network, we can try to calculate some metrics that will tell us a bit about the interactions we are seeing. We will first use bipartite to calculate all of the network metrics, and then go into depth for a few important metrics. 

```{r all the metrics}
networklevel(hummingbirds) #this may take a while to run
```
## Individual metrics: We can also call individual metrics, by using the networklevel() function, and specifying the index. For example
 
### Connectance, or the number of realized interactions over the number of potential interactions in the system. 
```{r connectance}
networklevel(hummingbirds, index = "connectance")
```
What does a connectance of 0.168 mean? It means that 16.8% of the total number of interactions are present in this network. 

### Nestedness, or how well the data forms into subsets of larger groups. We can use `weighted NODF`, which weights the interaction by how frequently we see it occur in the network. Higher values indicate higher nestedness. 

```{r}
networklevel(hummingbirds, index = "weighted NODF")
```
### Modularity, or when subgroups of species interactmore strongly with each over than with the rest of the network. Higher values indicate greater levels of modularity for the network. 
```{r}
networklevel(hummingbirds, index = "modularity")
```
resource: https://doi-org.ezproxy.lib.utexas.edu/10.1016/B978-0-12-409548-9.10564-0


