---
title: "Network Analysis"
Authours: Lili Benitez, Livia Raulinaitis, Sydney Jansen
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Network Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup}
library(ADANetworks)
```

## Load libraries, you may need to install the package bipartite
```{r load, message=FALSE, warning=FALSE}
#install.packages("bipartite")
library(tidyverse)
library(bipartite)
hummingbirds<- read.csv("EXAMPLE MATRIX_hummingbirdmatrix.csv", header = T) %>%
  column_to_rownames("X")
head(hummingbirds)
```

## Network Visualization: 
Lets look first at what this network looks like: we can do this using a couple of different functions, `plotweb` and `visweb`
```{r network visualization}
plotweb(sortweb(hummingbirds))
visweb(as.matrix(hummingbirds), labsize = 2)
```

## Calculating network metrics. We will first use bipartite to calculate all of the network metrics, and then go into depth for a few important metrics. 

```{r all the metrics}
networklevel(hummingbirds) #this may take a while to run
```
## Individual metrics: 
We can also call individual metrics, by using the networklevel() function, and specifying the index. For example: 
```{r connectance}
networklevel(hummingbirds, index = "connectance")
```


